Aligning marine species range data to better serve science and conservation

### 2016-03-27 case_study_mpa/3_mpa_spp_analysis.Rmd
pulling figures to update the supplemental info

``` {r combine all runs into one df}
mpa_all_runs <- mpa_klein_am2014 %>%
  mutate(trial = 'AM 2014 (50%)') %>%
  rename(cat = cat2014, pct = pct2014) %>%
  bind_rows(mpa_klein_am2015 %>%
    mutate(trial = 'AM 2015 (50%)') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(mpa_am2015_0 %>%
    mutate(trial = 'AM 2015 (0%)') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(mpa_am2015_iucn %>%
    mutate(trial = 'AM 2015 (0%) +\nIUCN 2015') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(iucn_spp_sum %>%
    mutate(trial = 'IUCN 2015') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  select(speciesid, cat, pct, trial)

>>>>

``` {r combine all runs into one df}
mpa_all_runs <- mpa_klein_am2014 %>%
  mutate(trial = 'AM 2014 (50%)') %>%
  rename(cat = cat2014, pct = pct2014) %>%
  bind_rows(mpa_klein_am2015 %>%
    mutate(trial = 'AM 2015 (50%)') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(mpa_am2015_0 %>%
    mutate(trial = 'AM 2015 (0%)') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(mpa_am2015_iucn %>%
    mutate(trial = 'AM 2015 (0%) +\nIUCN 2015') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(iucn_spp_sum %>%
    mutate(trial = 'IUCN 2015') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  select(speciesid, cat, pct, trial) %>%
  unique()

### 2016-03-27 ms_6_supplemental.Rmd
pulling figures to update the supplemental info

## Data processing

@@

AquaMaps data was provided by AquaMaps directly via ftp. Three SQL files were sent that contained (1) list of all `r nrow(spp_am)` species and their taxonomic information, (2) information for all (292,000?) half degree cells used for range mapping and (3) the probability of occurrence per cell for all `r nrow(spp_am)` species. These were extracted into .csv form using R Statistical Software.

@@

#### Possible figures

S1: sample maps from each quadrant

S2: risk by quadrant

S3: coral depth map

___Red List inclusion:___

<!--![](figures/bar_by_taxa_assessed.png)-->

![](figures/barchart_spp_risk_quads.png)

* Breaking down the quadrants by IUCN extinction risk categories (FIG 3c), we found little  with higher extinction risk tend to be better aligned between the two datasets, perhaps correlated to increased expert scrutiny.  Does higher perceived risk lead to increased attention, and thus better understanding of species distribution? Or conversely, does increased attention to species distribution reveal more species at risk?  Likely both mechanisms are at play on a case-by-case basis, depending on the species' taxon and region. _does this argument bear up to closer scrutiny? CR isn't dominated by Q1 any more_

## illustrative maps for different quadrants and different mechanistic problems

@@

@@

## AquaMaps: Effect of changing "presence" threshold on apparent distribution

### AquaMaps presence threshold analysis

### Figure 5 (a, b):

At a presence threshold of 40%, as used in the Ocean Health Index Species subgoal, the bulk of AquaMaps species suffer a significant decrease in represented range, and some species lose nearly their entire range.  Incrementing the presence threshold from 0.00 to 1.00 for the entire AquaMaps dataset, the shallow downward trend indicates a low but consistent sensitivity to threshold choice, with no surprising tradeoffs that could suggest an "optimal" threshold.  This pattern may not hold true for all subsets of AquaMaps species, however, whether subsetting by taxa or by georegion.

### MPA gap analysis - explanation

>>>>

## Data processing - should this go in methods???

@@

AquaMaps data was provided by AquaMaps directly via ftp. Three SQL files were sent that contained (1) list of all `r nrow(spp_am)` species and their taxonomic information, (2) information for all (259,200?) half degree cells used for range mapping and (3) the probability of occurrence per cell for all `r nrow(spp_am)` species. These were extracted into .csv form using R Statistical Software.

@@

## S1: sample maps from each quadrant

## S2: risk by quadrant

![](figures/si_barchart_spp_risk_quads.png)

* Breaking down the quadrants by IUCN extinction risk categories (Fig. S2), we found little support for our hypothesis that maps for species with higher extinction risk tend to be better aligned between the two datasets, perhaps correlated to increased expert scrutiny.  Does higher perceived risk lead to increased attention, and thus better understanding of species distribution? Or conversely, does increased attention to species distribution reveal more species at risk?  Likely both mechanisms are at play on a case-by-case basis, depending on the species' taxon and region. _does this argument bear up to closer scrutiny? CR isn't dominated by Q1 any more_

## S3: coral depth map

@@

## S4: AquaMaps threshold examination

S4a:

S4b:

At a presence threshold of 40%, as used in the Ocean Health Index Species subgoal, the bulk of AquaMaps species suffer a significant decrease in represented range, and some species lose nearly their entire range.  Incrementing the presence threshold from 0.00 to 1.00 for the entire AquaMaps dataset, the shallow downward trend indicates a low but consistent sensitivity to threshold choice, with no surprising tradeoffs that could suggest an "optimal" threshold.

## MPA gap analysis - explanation

<!--
___Red List inclusion:___

![](figures/bar_by_taxa_assessed.png)-->

@@
